---
title: Reaching for the stars
author: ~
date: '2019-06-11'
slug: reaching-for-the-stars
categories: []
tags: []
keywords:
  - tech
---

```{r, include=FALSE}
library(tidyverse)
library(RCurl)
library(naniar)
```

## Get the data from github

I couldn't be bothered downloading the data and saving it on my laptop (it fills up so quickly!). Also, I have been using github more lately and I figured I should learn how to download the data directly from github. I first tried to read in the file using the link that you get when you click on the .csv file.
```{r}
wrong_df <- read_csv('https://github.com/rfordatascience/tidytuesday/blob/master/data/2019/2019-06-11/meteorites.csv')
head(wrong_df)
```

That was not what I wanted! It turns out you need to click on 'view raw' and use that link. It will have `raw.githubusercontent` in the url.
```{r}
df <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-11/meteorites.csv")
head(df)
```
Now we are talking!

## Now look at the data

I wanted to try some new packages and decided to explore the `naniar` package.



```{r}
world<-data("World")

tm_shape(World) +
    tm_polygons("HPI")
```


```{r}
(bb_world = st_bbox(world)) # the world's bounds


df_no_na <- df %>% 
  filter(!is.na(lat))
random_points <- df_no_na %>% 
  st_as_sf(coords = c("long", "lat")) %>% # set coordinates
  st_set_crs(4326) # set geographic CRS


world_random = world[random_points, ]
nrow(world_random)
#> [1] 4
random_joined = st_intersects(random_points, world["name_long"])
```

