---
title: Regressions using completly fake data!
author: ~
date: '2018-01-13'
slug: fake-data
categories:
  - R
tags:
  - simstudy
  - texreg
  - tableone
  - purrr
---


```{r packages, include=FALSE}
library("tidyverse")
library("survival")
library("simstudy")
library("texreg")
library("tableone")
```

# Why can't I just have all of the data?

I have been updating some code that uses data that I don't have access to. I used one of the data sets from the `survival` package but it annoyed me that the variable names were not what I wanted them to be and I couldn't be bothered changing them. I was killing time on a train the other day and discovered the `simstudy` package and decided to give it a go. It worked beautifully and made updating my code so much easier. It also made me realise how much R I have learnt in the last year so I wanted to show the difference between the old and the new code with you.   

# Creating data
Start by defining the variables you want in your data set and create the survival part of the data and before creating the full data set.
```{r}
def <- defData(varname = "v1_6", dist = "binary", formula = 0.6, id = "idnum", link = "logit")
def <- defData(def, varname = "v1_5", dist = "uniformInt", formula = "18;110")
def <- defData(def, varname = "v2_1_1", dist = "binary", formula = 0.5)
def <- defData(def, varname = "v2_1_2", dist = "binary", formula = 0.6)
def <- defData(def, varname = "v2_1_3", dist = "binary", formula = 0.5)
def <- defData(def, varname = "v2_1_4", dist = "binary", formula = 0.55)
def <- defData(def, varname = "v2_1_5", dist = "binary", formula = 0.5)
def <- defData(def, varname = "v2_5", dist = "binary", formula = 0.5)
def <- defData(def, varname = "v7_4", dist = "binary", formula = 0.5)
def <- defData(def, varname = "event", dist = "binary", formula = 0.1)
def <- defData(def, varname = "died", dist = "binary", formula = 0.5)
def <- defData(def, varname = "NIHSS_15", dist = "uniformInt", formula = "0;42", link="identity")
def <- defData(def, varname = "age_group", dist = "categorical", formula = "0.5;0.2;0.3", link="identity")
# Define survival data part
sdef <- defSurv(varname = "survTime", formula = "v1_5/2", scale = 20, 
                shape = "v2_5*1 + (1-v2_5)*1.5")
sdef <- defSurv(sdef, varname = "censorTime", scale = 80, shape = 1)
# Create data
df <- genData(5000, def)
df <- genSurv(df, sdef)
```

What a lovely data set I created!

# Descriptive statistics

Now that I have data I want to compare the group that experienced the event to the group that didn't. I use CreateTableOne from the `tableone` package. I need to specify which variables I want to include (or all variables will be included) and which ones I want to be factors (will be shown as %) or nonnormal variables (will be shown as median). 
```{r}
table_var<- c("v1_5", "v1_6", "v2_1_1", "v2_1_2", "v2_1_3", "v2_1_4", "v2_1_5", "v2_5","v7_4", "age_group")
median_var <- median_var<- c("v1_5", "NIHSS_15", "v2_2", "v7_4")
factor_var <- c("v1_6","v2_5","v2_1_1", "v2_1_2", "v2_1_3", "v2_1_4", "v2_1_5","age_group")

table1<- CreateTableOne(table_var, data = df,  strata = "event", factorVars = factor_var,
                        testNonNormal = kruskal.test, argsNonNormal = median_var)
table1_print<-print(table1, nonnormal = median_var, contDigits=0, 
                 exact = "LOC",  printToggle = TRUE, factorVars = factor_var)

```

You might want to save the table in a different format such as a csv file. This code should help you do that. Note that the print options are slightly different. 
```{r}

table1_print <- print(table1, nonnormal = median_var, contDigits=0, 
                 exact = "LOC", quote = FALSE, noSpaces = TRUE, printToggle = FALSE, factorVars = factor_var)
## Save to a CSV file
write.csv(table1_print, file = "my_table1.csv")

```


# Regressions
 
Time to move on to the regressions! I wrote the first version of this code about 9 months ago and it looked a bit like this but with more models:

```{r silly way, eval=FALSE}

surv_data <- Surv(time=df$survTime, event=df$event, type="right")

coxph.fit.v1.6 <- coxph(surv_data ~ v1_6 , data=df, method="breslow")
coxph.fit.v1.5 <- coxph(surv_data ~ v1_5 , data=df, method="breslow")
coxph.fit.v2.5 <- coxph(surv_data ~ v2_5 , data=df, method="breslow")

                            
```

As you can imagine this made my code really long and included a lot of copying and pasting. 

As I learnt more and more R (and discovered the power of `purrr` after an R ladies event where Charlotte Wickham presented) I decided to update the code before giving to the person who will run it. I start by putting all of my predictors in a vector.  I specify all the predictors I care about and  use `map` to fit the model for each one. 

```{r}

predictors <- c("v1_5", "v1_6", "v2_1_1", "v2_1_2", "v2_1_3", "v2_1_4", "v2_1_5", "v2_5","v7_4", "age_group")

surv_data <- Surv(time=df$survTime, event=df$event, type="right")

unadjusted <- map(predictors, ~coxph(as.formula(paste("surv_data", .x, sep="~")), data=df))

```
 
I refuse to copy numbers off the screen because I know I will get it wrong so I use the `texreg` package to create nice regression tables. I need to override the coefficients as I am interested in exp(coef) and this is how I did it 9 months ago. I spent most of my time terrified that I would accidentally use coefficients from the wrong model and adding new models took forever. 
```{r, eval=FALSE}
screenreg(list(coxph.fit.v1.6, coxph.fit.v1.5, coxph.fit.v2.5 ),  
        single.row= TRUE , ci.force= TRUE, include.aic = TRUE, include.rsquared =  TRUE, include.maxrs= FALSE,
        ci.test = 1,
        override.coef = list(summary(coxph.fit.v1.6)$coef[,2],
                             summary(coxph.fit.v1.5)$coef[,2],
                             summary(coxph.fit.v2.5)$coef[,2])
        )
```


Once again the `map` function saved me. I create lists with the coefficients and upper and lower confidence intervals. This way I don't have to worry about using the wrong coefficients and it is real easy to add more models. 
 
```{r}
unadjusted_coef <- map(unadjusted, summary ) %>% 
  map("coefficients") %>% 
  map(as.data.frame) %>% 
  map("exp(coef)")

unadjusted_lower_conf <- map(unadjusted, summary) %>% 
  map("conf.int") %>% 
  map(as.data.frame) %>% 
  map("lower .95")

unadjusted_upper_conf <- map(unadjusted, summary) %>% 
  map("conf.int") %>% 
  map(as.data.frame) %>% 
  map("upper .95")

screenreg(unadjusted, single.row= FALSE, ci.force= TRUE, ci.test=1,
          override.coef = unadjusted_coef,
          override.ci.low=unadjusted_lower_conf,
          overide.ci.up=unadjusted_upper_conf)
```


It is now super easy to add/remove variables to my models. I already know that I want to include two variables - v1_5 and v1_6 - in my full model so I add them to the model. I remove v1_5 from my predictors or I will have two models for just v1_5 and v1_6.  

```{r}

predictors <- c("v1_6", "v2_1_1", "v2_1_2", "v2_1_3", "v2_1_4", "v2_1_5", "v2_5","v7_4", "age_group")

adjusted <- map(predictors, ~coxph(as.formula(paste("surv_data ~ v1_5 + v1_6", .x, sep="+")),
                                   data=df))

adjusted_coef <- map(adjusted, summary ) %>% 
  map("coefficients") %>% 
  map(as.data.frame) %>% 
  map("exp(coef)")

adjusted_lower_conf <- map(adjusted, summary) %>% 
  map("conf.int") %>% 
  map(as.data.frame) %>% 
  map("lower .95")
  
adjusted_upper_conf <- map(adjusted, summary) %>% 
  map("conf.int") %>% 
  map(as.data.frame) %>% 
  map("upper .95")

screenreg(adjusted, single.row= FALSE, ci.force= TRUE, ci.test=1, 
          override.coef = adjusted_coef,
          override.ci.low=adjusted_lower_conf,
          overide.ci.up=adjusted_upper_conf)
```


If you want to save the models in a word document this is how to do it. 

```{r}

htmlreg(adjusted, single.row= FALSE, ci.force= TRUE, ci.test=1, type="html", file="My_regressions.doc",
          override.coef = adjusted_coef,
          override.ci.low=adjusted_lower_conf,
          overide.ci.up=adjusted_upper_conf)
 
```

# Packages I used
```{r , eval=FALSE}
library("tidyverse")
library("survival")
library("simstudy")
library("texreg")
library("tableone")
```
